Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-04 robdml.tex
page 1

Existing defense methods for DML~\cite{advrank,robrank} are adversarial
training-based, inspired by Madry's \emph{min-max} adversarial
training~\cite{madry} because it is one of the most effective methods for deep
neural network classifiers.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse, the existing DML defenses refrain
from directly adopting such paradigm, but instead use some indirect and
relatively weak objectives to increase the loss value to a certain level.
%
As a result, the model is not learning from a strong adversary, and suffers
from low efficiency in gaining robustness within limited time. 
%
Furthermore, adversarial training already suffers from significant training
cost (compared to the normal training process).
%
Hence, efficiency in gaining adversarial robustness is an inevitable major
issue for DML defense methods.

@@

As suggested by previous works, an appropriate adversary for the inner
\emph{maximization} problem should increase the loss to a certain point between
that of benign examples (\ie, unperturbed examples) and the theoretical upper
bound.
%
What if this point is reached by a strong adversary, directly and
efficiently?
%
Besides, as the mathematical expectation of loss is also greatly influenced by
the triplet sampling strategy, we conjecture that the triplet sampling strategy
and triplet ``hardness'' have a key impact for adversarial training.

>>>>

Existing defense methods~\cite{advrank,robrank} are adversarial training-based,
inspired by Madry's \emph{min-max} adversarial training~\cite{madry} because it
is consistently one of the most effective methods for classification task.
%
Specifically, Madry's method involves a inner problem to \emph{maximize} the
loss by perturbing the inputs into adversarial examples, and an outer problem
to \emph{minimize} the loss by updating the model parameters.
%
However, in order to avoid model collapse due to excessively hard examples, the
existing defenses refrain from directly adopting such min-max paradigm, but
instead replace the inner problem to indirectly increase the loss value to a
certain level, which suffers from low efficiency and weak adversary (and hence
weak robustness).
%
Since training cost is already a serious issue of adversarial training, the
efficiency in gaining higher adversarial robustness under a fixed budget is
inevitable and important for DML defense.

@@

Previous works suggest that an appropriate adversary for the inner
\emph{maximization} problem should increase the loss to an ``intermediate''
point between that of benign examples (\ie, unperturbed examples) and the
theoretical upper bound.
%
In this paper, we argue such point can be reached by a direct and efficient
adversary.
%
Besides, we conjecture the triplet sampling strategy has a key impact in
adversarial training, because it is also able to greatly influence the
mathematical expectation of loss even without adversarial attack.
